Evaluating `Graphical Perception' with CNNs

### 2018-03-31 PAPER/paper.tex
Going through Section 4.

%% Abstract section.
\abstract{%Convolutional neural networks are being successfully used for image understanding and object recognition tasks while regularly outperforming humans. Despite their tremendous success, little is known perceptional capabilities. In there Graphical Perception paper from 1984, Cleveland and McGill define elementary perceptual tasks that let people extract quantitative information from visualizations and measure human perception of different visual encodings with in depth user studies. 
%We replicate the experimental setup of Cleveland and McGill and evaluate the perceptional capabilities of four modern classifiers based on neural networks. We systematically test how the classifers perform on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment which compares pie and bar charts, c) the position-length experiment which compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues help to perceive information. We present the results of these experiments to foster the understanding of how such classifiers can be successfully applied to data visualizations. On this journey, we also study how the feed-forward neural networks obey Weber's law which defines the proportional relation between perceivable information and distribution size. We further introduce a ranking of elementary visual encodings targeted towards modern deep neural networks and derive practical evidence for properties of convolutional neural networks such as the translation invariance.  
Convolutional neural networks can successfully perform many computer vision tasks on images, and their learned representations are often said to mimic the early layers of the visual cortex. But can CNNs understand graphical perception for visualization? We investigate this question by reproducing Cleveland and McGill's seminal 1984 experiments, which measured human perception efficiency of different visual encodings and defined elementary perceptual tasks for visualization. We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment that compares pie charts to bar charts, c) the position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues aid perception. We also study how feed-forward neural networks obey Weber's law, which defines the proportional relation between perceivable information and distribution density. We present the results of these experiments to foster the understanding of how CNN classifiers succeed and fail when applied to data visualizations.
} % end of abstract

>>>>

%% Abstract section.
\abstract{
Convolutional neural networks can successfully perform many computer vision tasks on images, and their learned representations are often said to mimic the early layers of the visual cortex. But can CNNs understand graphical perception for visualization? We investigate this question by reproducing Cleveland and McGill's seminal 1984 experiments, which measured human perception efficiency of different visual encodings and defined elementary perceptual tasks for visualization. We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment that compares pie charts to bar charts, c) the position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues aid perception. We also study how feed-forward neural networks obey Weber's law, which defines the proportional relation between perceivable information and distribution density. We present the results of these experiments to foster the understanding of how CNN classifiers succeed and fail when applied to data visualizations.
} % end of abstract
